Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-02-26 article-models-tt.tex
small impro
resolving issue #23

\subsection{Extension on terms}\label{extension-on-terms}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we obtain a term $\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}$. This operation
of weakening on terms is like the pairing operation and like with the pairing
operation, extension on terms tells what the terms of an extended family are. 
However, it is handled
much stricter than $\smsym$-types, 
for it has a judgmental inverse and it is subject to various compatibility
rules.
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\pts{w}}{\edg{w}}}{w}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\pts{(\ctxext{x}{u})}}{x}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\edg{(\ctxext{x}{u})}}{u}}
\end{align}

@@

\subsubsection{Identity functions}
Without a rule explicitly asserting the existence of an identity morphism we don't
get one, hence we do that here. The identity morphism is a term which introduced
in ordinary type theory via the variable rule. The variable rule is a bit more
general: it asserts that
\begin{equation*}
\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
\end{equation*}
for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
we get the projections from the identity morphisms together with weakening. We
already have weakening, so here it suffices to introduce the identity morphisms.
They are determined by their behavior with respect to substitution combined with
weakening. The identity functions will also be subject to compatibility rules.
\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}\\
& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
\end{align}

>>>>

\subsection{Extension on terms}\label{extension-on-terms}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we obtain a term $\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}$. This operation
of weakening on terms is like the pairing operation and like with the pairing
operation, extension on terms tells what the terms of an extended family are:
every term of an extended family decomposes into the components of a pair. 
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}
& & \inference{\jtermeq{\Gamma}{A}{x}{x'}\quad\jtermeq{\Gamma}{\subst{x}{P}}{u}{u'}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}{\ctxext{x'}{u'}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}
& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{A}{\pts{w}}{\pts{w'}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}
& & \inference{\jtermeq{\Gamma}{\ctxext{A}{P}}{w}{w'}}{\jtermeq{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}{\edg{w'}}}
\end{align}
However, this pairing is handled
much stricter than $\smsym$-types, 
because the decomposition is a judgmental inverse (and it will be subject to various compatibility
rules).
\begin{align}
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\pts{w}}{\edg{w}}}{w}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\pts{(\ctxext{x}{u})}}{x}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\edg{(\ctxext{x}{u})}}{u}}
\end{align}

@@

\subsubsection{Identity functions}
Without a rule explicitly asserting the existence of an identity morphism we don't
get one, hence we do that here. The identity morphism is a term which introduced
in ordinary type theory via the variable rule. The variable rule is a bit more
general: it asserts that
\begin{equation*}
\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
\end{equation*}
for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
we get the projections from the identity morphisms together with weakening. We
already have weakening, so here it suffices to introduce the identity morphisms.
\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jterm{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
& & \inference{\jfameq{\Gamma}{A}{A'}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
\end{align}
Identity functions are determined by their behavior with respect to substitution combined with
weakening. The identity functions will also be subject to compatibility rules.
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}\\
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}} \\
& \inference{\jterm{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
\end{align}
